---
title: "turbokit_tidymath"
output: rmarkdown::html_vignette
author: "Donald Seinen"
vignette: >
  %\VignetteIndexEntry{turbokit_tidymath}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(turbokit)
```

In this vignette, *tidymath* is introduced, covering the following topics:

-   What is tidymath?

-   Syntax

    -   single functions

    -   chaining functions

-   

[tidy](https://www.jstatsoft.org/article/view/v059i10)

Let's combine them. The turbokit package provides quality of life features that makes existing systems and the tidy design philosophy easily accessible, more consistent and speedier to use.

Tidymath is one such feature, which seeks to provide a way to "vectorise" code input for frequently used functions, which is performing the same operation on a set of values.

How many functions are typed, how many backspaces hit every day, every month, every year, per career, per R user?

> Do not Repeat Yourself (Hunt & Davis, 1999)

Adults on average, can read about 270 words per minute

The same adult would type about 40-60 words per minute.

## What is tidymath?

*Tidymath* is a translation mechanism for an initialism description of frequently used functions, their interactions and origins. It is a general approach to describe function names in abbreviated format, and uses a set of symbols to identify package names, placements of function and arguments, syntax and sequence of function calls and operators.

Tidymath can be seen as the crankshaft of a car: it translates

An initialism is an abbreviation formed from initial letters, in the case of tidymath, the initials of each component of a function name. For example, the tidyselect function `starts_with()` maps to `sw` .

Tidymath powers quick input by reducing typing of functions to performing the same operation on a set of values. The values in *tidymath* are the names of functions in abbreviated form.

Consider the following piece of legacy code:

```{r BaseR, eval=FALSE}
transform(iris[which(iris$Species == "versicolor"),], new = Sepal.Length*2)
```

In baseR such a call would take roughly 76 keystrokes, given that there are no mistakes.

```{r Tidyverse, eval=FALSE}
iris %>%
    select(starts_with("Sepal")) %>%
    filter(str_detect())
    mutate(new = Sepal.Length * 2)
```

The tidyverse packages provide an alternative way to write code in a nice format

```{r, eval=FALSE}
iris %>%
    select(starts_with("Sepal"), Species) %>%
    filter(str_detect(Species, "v")) %>%
    group_by(Species) %>%
    summarise(across(everything(), mean)) %>%
    ggplot(aes(Sepal.Length, Sepal.Width)) +
    geom_density_2d_filled()
```

What the current design philosophy does not do, however, is heavily reduce the keypresses required to arrive at this code. The above example takes roughly 75 keystrokes.

Turbokit provides two new ways to input the code in tidy format.

### method 1: shortcuts

```{r, eval=FALSE}
"~f>m" %>>%.
s
```

which returns

```{r, eval=FALSE}
snippet s
	${1:mydata} %>%
	  filter(${2:var}) %>%
	  mutate(${3:var})
```

turbokit: 53 key presses

The integrated development environment (IDE) RStudio has many systems in place that aid the user to reduce cognitive and physical workload when using R.

1.  Functions

2.  Objects

3.  Code autocompletion

4.  Code snippets

5.  Addins

6.  Keyboard shortcuts

Inspiration for tidymath and superpipe comes from several sources.

-   Recipes and workflow packages

-   Tidy

-   Plotmath

-   Observation

## Defining Tidymath

Definition

:   **tidymath** is a shorthand notation form for widely used R functions. It uses first letters of snake case functions to construct a complex pipeline of function calls.

*Example*: `"~s*sw>m*2sd>z*a*e>p>1gv>1syc"`, which is equivalent to

```{r, eval=FALSE}
mydata %>%
  select(starts_with()) %>%
  mutate(str_detect()) %>%
  summarise(across(everything())) %>%
  ggplot(aes()) +
  geom_violin() +
  scale_y_continuous()
```

## Tidymath terminology

Defintions

:   The **initial term** ( \~ ) denotes the system of functions to use. Currently supported are the tidyverse ( \~ ) and tidymodels.

:   An **interaction term** ( \* ) denotes that the function on the right is placed within the parenthesis of the function on the left.

:   A **package term** ( digit ) denotes the package in which the function lives, exceptions are the dplyr and tidyselect packages, for which no package term is implemented.

:   An **abbreviation** describes a compound function by its first snake case letters.

:   A **function term** ( letter ) denotes a part or the whole of an abbreviation for a function.

# Design origins

There are 4 events that led to the implementation of turbokit.

1.  A catastrophic typo by its author on a Sunday afternoon just after a cup of coffee. `group_By()+`

    Wrong operator, capital letter after underscore, no space after ). 5 backspaces, 5 keypresses.

2.  Writing code is a bottleneck in the data science process.

    Adults read about 270-300 words per minute, but type only about 40-70 words per minute given *fluency* in a language. Now extend this to non-native speakers and different keyboard layouts or scripts. An example is Chinese, where the *Shift* key is often linked to a change in input styles. Punctuation marks and input often is not identical to what R expects and recognizes, for example the parenthesis ）verses ).

3.  Realisation that much of a function name is redundant.

> It deosn't mttaer in waht oredr the ltteers in a wrod are, hmuans can udnertsand it.

If we then do not write for computers, nor place letters in the correct order for ourselves, who or what are we writing so much text for?

`starts_with()` `startsWith()`

4.  Existing systems are wonky and inefficient, but powerful.

5.  Best practices evolve faster than
