---
title: "turbokit_tidymath"
output: rmarkdown::html_vignette
author: "Donald Seinen"
vignette: >
  %\VignetteIndexEntry{turbokit_tidymath}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(turbokit)
```

In this vignette, *tidymath* is introduced, covering the following topics:

-   What is tidymath?

    -   Definition

    -   Superpipe - translating tidymath

    -   Dynamic snippets - inserting the result

-   What are dynamic snippets?

    -   Problem definition

    -   Design philosophy

-   Syntax

    -   single functions

        -   Exceptions

    -   chaining functions

-   

# What is tidymath?

The turbokit package provides quality of life features that makes existing systems and the tidy design philosophy easily accessible, more consistent and speedier to use. Tidymath is one such feature. *Tidymath* is an initialism description of frequently used functions, their interactions and origins. It is a general approach to describe function names in abbreviated format

Abbreviations used in tidymath adhere to [tidy](https://www.jstatsoft.org/article/view/v059i10) principles, that is, each abbreviation is structured and constructed in the same way, it does not matter if the function it describes is snake case, camel case or some other notation form. Using a metaphor, tidymath is "vectorising" function names allowing for performing the same operation on vastly different function naming conventions.

```{r, eval=FALSE}
# Tidymath
"~s*sw>m*3fr>g>p>1gh"

```

# The superpipe - translating tidymath

The superpipe `%>>%` is the translation mechanism that converts tidymath into usable complex pipes.

```{r, eval=FALSE}
# Tidymath + superpipe
"~s*sw>m>g>p>1gh" %>>%.

# Describes
mydata %>%
    select(starts_with()) %>%
    mutate() %>%
    group_by() %>%
    ggplot(aes()) +
    geom_histogram()

```

## Dynamic snippets - inserting tidymath + superpipe

The superpipe serves another purpose. In addition to translating tidymath, it converts the outcome to a *dynamic snippet*. Dynamic snippets are [snippets](%5B%3Chttps://support.rstudio.com/hc/en-us/articles/204463668-Code-Snippets%3E)](<https://support.rstudio.com/hc/en-us/articles/204463668-Code-Snippets>)), text macros, modified in place according to a singular naming convention for one-time use. This differs from regular snippets in that those usually are defined in advance to simplify recurring coding tasks.

```{r, eval=FALSE}
# Tidymath + superpipe
"~s*sw>m*3fr>g>p>1gh" %>>%.

# Translates to
${1}:mydata %>%
    select(starts_with("${2}:charmatch")) %>%
    mutate(${3:new}) %>%
    group_by(${4:var}) %>%
    ggplot(aes(${5:param})) +
    geom_histogram()
    

```

The last piece of the puzzle that the superpipe does, is to insert this dynamic snippet at the cursor location. The user can then call a custom snippet, named `s` by writing an s and hitting the `TAB` key. For Windows users, an additional package setting is available, `getOption("turbokit-autoinsert")` to automatically insert the snippet after its creation. Alternatively, users might want to use the insertSnippet shortcut.

# What is complex input?

Complex input is a one-off execution of tidymath through an interface. This interface does not break the dynamic snippet, and can be used to append pipes after they have been constructed.

## Problem definition

How many functions are typed, how many backspaces hit every day, every month, every year, per career, per R user? What is the total redundant hand travel distance? Once asked, a problem becomes clear.

> Do not Repeat Yourself (Hunt & Davis, 1999)

Adults on average, can read about 270 words per minute

The same adult would type about 40-60 words per minute.

## What is tidymath?

*Tidymath* is a translation mechanism for an initialism description of frequently used functions, their interactions and origins. It is a general approach to describe function names in abbreviated format, and uses a set of symbols to identify package names, placements of function and arguments, syntax and sequence of function calls and operators.

Tidymath can be seen as the crankshaft of a car: it translates

An initialism is an abbreviation formed from initial letters, in the case of tidymath, the initials of each component of a function name. For example, the tidyselect function `starts_with()` maps to `sw` .

Tidymath powers quick input by reducing typing of functions to performing the same operation on a set of values. The values in *tidymath* are the names of functions in abbreviated form.

Consider the following piece of legacy code:

```{r BaseR, eval=FALSE}
transform(iris[which(iris$Species == "versicolor"),], new = Sepal.Length*2)
```

In baseR such a call would take roughly 76 keystrokes, given that there are no mistakes.

```{r Tidyverse, eval=FALSE}
iris %>%
    select(starts_with("Sepal")) %>%
    filter(str_detect())
    mutate(new = Sepal.Length * 2)
```

The tidyverse packages provide an alternative way to write code in a nice format

```{r, eval=FALSE}
iris %>%
    select(starts_with("Sepal"), Species) %>%
    filter(str_detect(Species, "v")) %>%
    group_by(Species) %>%
    summarise(across(everything(), mean)) %>%
    ggplot(aes(Sepal.Length, Sepal.Width)) +
    geom_density_2d_filled()
```

What the current design philosophy does not do, however, is heavily reduce the keypresses required to arrive at this code. The above example takes roughly 75 keystrokes.

Turbokit provides two new ways to input the code in tidy format.

### method 1: shortcuts

```{r, eval=FALSE}
"~f>m" %>>%.
s
```

which returns

```{r, eval=FALSE}
snippet s
	${1:mydata} %>%
	  filter(${2:var}) %>%
	  mutate(${3:var})
```

turbokit: 53 key presses

The integrated development environment (IDE) RStudio has many systems in place that aid the user to reduce cognitive and physical workload when using R.

1.  Functions

2.  Objects

3.  Code autocompletion

4.  Code snippets

5.  Addins

6.  Keyboard shortcuts

Inspiration for tidymath and superpipe comes from several sources.

-   Recipes and workflow packages

-   Tidy

-   Plotmath

-   Observation

## Defining Tidymath

Definition

:   **tidymath** is a shorthand notation form for widely used R functions. It uses first letters of snake case functions to construct a complex pipeline of function calls.

*Example*: `"~s*sw>m*2sd>z*a*e>p>1gv>1syc"`, which is equivalent to

```{r, eval=FALSE}
mydata %>%
  select(starts_with()) %>%
  mutate(str_detect()) %>%
  summarise(across(everything())) %>%
  ggplot(aes()) +
  geom_violin() +
  scale_y_continuous()
```

## Tidymath terminology

Defintions

:   The **initial term** ( \~ ) denotes the system of functions to use. Currently supported are the tidyverse ( \~ ) and tidymodels.

:   An **interaction term** ( \* ) denotes that the function on the right is placed within the parenthesis of the function on the left.

:   A **package term** ( digit ) denotes the package in which the function lives, exceptions are the dplyr and tidyselect packages, for which no package term is implemented.

:   An **abbreviation** describes a compound function by its first snake case letters.

:   A **function term** ( letter ) denotes a part or the whole of an abbreviation for a function.

# Design origins

There are 4 events that led to the implementation of turbokit.

1.  A catastrophic typo by its author on a Sunday afternoon just after a cup of coffee. `group_By()+`

    Wrong operator, capital letter after underscore, no space after ). 5 backspaces, 5 keypresses.

2.  Writing code is a bottleneck in the data science process.

    Adults read about 270-300 words per minute, but type only about 40-70 words per minute given *fluency* in a language. Now extend this to non-native speakers and different keyboard layouts or scripts. An example is Chinese, where the *Shift* key is often linked to a change in input styles. Punctuation marks and input often is not identical to what R expects and recognizes, for example the parenthesis ）verses ).

3.  Realisation that much of a function name is redundant.

> It deosn't mttaer in waht oredr the ltteers in a wrod are, hmuans can udnertsand it.

If we then do not write for computers, nor place letters in the correct order for ourselves, who or what are we writing so much text for?

`starts_with()` `startsWith()`

4.  Existing systems are wonky and inefficient, but powerful.

5.  Best practices evolve faster than
