---
title: "turbokit"
output: rmarkdown::html_vignette
description: >
  Start here if this is your first time using turbokit. You'll learn how to setup, the       package, its basic philosophy, the most important code manipulation actions.
vignette: >
  %\VignetteIndexEntry{turbokit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette explains in detail each of these features.

-   Package summary

-   Turbokit Introduction

    -   Why this package

    -   Development roadmap

-   Setup

    -   Shortcuts

    -   Options

-   Turbokit design philosophy

# Package Summary

Turbokit introduces quality of life features reducing redundancy in day-to-day coding tasks in R and increasing the accessibility of stylish R code. It does so by re-thinking how frequently used functions are written by users, minimizing the number of key presses and reducing time spent on formatting tasks. Turbokit leverages [snippets](https://support.rstudio.com/hc/en-us/articles/204463668-Code-Snippets), [infix operators](https://cran.r-project.org/doc/manuals/R-lang.html#Operators), [custom shortcuts](https://support.rstudio.com/hc/en-us/articles/206382178-Customizing-Keyboard-Shortcuts), fuzzy matching, [addins](https://rstudio.github.io/rstudioaddins/) and [the tidyverse manifesto](https://cran.r-project.org/web/packages/tidyverse/vignettes/manifesto.html) to do so.

Turbokit includes the following features:

-   *Smart pipe:* A single button press to insert `%>%` or `+` based on document context.

see `vignette("turbo")`

-   Shortcut button for a curated list of (1500+) functions.

-   Shortcut buttons for extremely frequent functions.

see `vignette("chaining")`

-   *Pipemath*: Shorthand notation for writing chains of functions in a pipe format.

-   *Superpipe*: `%>>%` operator to convert *pipemath* instructions to a full-length snippet.

-   *Dynamic snippets:* adaptive snippets of formatted complex pipes described by *pipemath* and constructed by the *superpipe*.

-   *Quasi-comments*: Non-evaluated syntax highlighted R code.

# Turbokit Introduction

## Why turbokit?

A stigma yet clings to R - It is hard to use.

Historically, naming conventions used in R packages were [problematic](https://journal.r-project.org/archive/2012-2/RJournal_2012-2_Baaaath.pdf). Dots, camelCase, snake_case and function parameter names, many were unpredictable. This forced users to apply trial-and-error style probing to find out or remember what the function name is in a given package. Typos and deviation from that what matters: data science.

### Code input

There is however a bottleneck. Inputting code takes time, and this time cost increases the longer or harder to remember a function name is. Systems are in place to aid the user, but often come with hurdles to overcome or drawbacks when used.

+----------------+-------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| System         | Benefit                                                                                                                                         | Drawback                                                                                                                                                                                     |
+================+=================================================================================================================================================+==============================================================================================================================================================================================+
| Autocompletion | Visual cues and menuing for fuzzy matches, shorthand function notation. Covers all possible functions. Prompt speed and wait time customizable. | Packages have to be attached. to be searched. Output differs depending on typing speed. User has to slow down to be fast for most functions. Unpredictable position of match in search list. |
+----------------+-------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Shortcuts      | One-stop-shop for complicated input at a button press.                                                                                          | Requires package addin (long setup time) and shortcut overload as each button is tied to a unique key.                                                                                       |
+----------------+-------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Snippet        | Insert arbitrarily long code chunks, and jump from location to location.                                                                        | Requires special syntax and menuing to set up.                                                                                                                                               |
+----------------+-------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

While powerful in their own right, the systems available do not play very well with each other. There thus exists a gap - fast paced, predictable output for complicated functions that does not require complex syntax.

The turbokit package provides quality of life features that makes existing systems and the tidy design philosophy easily accessible, more consistent and speedier to use.

### Script navigation

Another bottleneck is script navigation and formatting code. Ever typed a wrong pipe operator? Forgot to insert library calls? Typed a capital letter instead of a small letter? Forgot the shortcut to move to the end of the line, resulting in abuse of the direction keys? Turbokit helps with this by providing some formatting tools.

# Setup

Turbokit is somewhat different from most packages, as many of its features are available without the need to call `library(turbokit)`. However, to use turbokit to its full potential, the user has to invest some time in setting up turbokit to their preference.

## Shortcuts

Several turbokit functions are exported as addin that can be bound to a shortcut key. Here, shortcuts are meant to be used as a quick insertion of a particular verb that work well with the smart pipe. These shortcuts will place the cursor between the brackets of the verb inserted at the cursor location. The smart pipe, when called, will relocate the cursor to the end of the line, and insert an operator based on the context of the pipe at that point.

Turbokit shortcut addins all start with "Turbo".

***Tools \> Addins \> Browse Addins \> Keyboard Shortcuts \> Turbo \_\_\_***

The following table is a suggestion for mapping turbokit shortcut keys. Some of these recommended shortcuts may overlap with default shortcuts. Recommended shortcuts are such that hand travel distance is minimized and shortcuts have a link to the function name.

Naturally, users are encouraged experiment with the shortcut setup to find their favorite setup that works best for them.

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
tabl <- "
| Function       | Shortcut 
|---------------|:-------------:|
| switch pipe      | *ctrl + shift + ,*  |
| switch mode      | *ctrl + shift + right*  |
| insert pipe      | *ctrl + shift + .*  |
| insert superpipe      | *ctrl + shift + `*   |
| turbo | *ctrl + shift + /*       |
| select       | *ctrl + shift + s*    |
| filter    | *ctrl + shift + f*       |
| mutate | *ctrl + shift + m*        |
| summarise      | *ctrl + shift + z*    |
| everything      | *ctrl + shift + e*    |
| across    | *ctrl + shift + a*       |
| starts_with | *ctrl + shift + 1*        |
| ends_with | *ctrl + shift + 2*        |
"
cat(tabl)
```

Besides the two pipe operators, all of the output of the insert\_ functions, can be achieved through pipemath and superpipe combination as well as the turbo function. It is thus recommended to only bind those functions that the user very frequently uses.

## Options

Turbokit has several options that can change the output of its operators and functions to the users desire or requirement. As turbokit deals with snippets and shortcuts, it inevitably has to deal with different operating systems and use cases.

Changes to options are not stored permanently, and must be rerun after restarting an R session. A way to permanently can be done by editing an Rprofile file.

`usethis::edit_r_profile() or  file.edit(file.path("~", ".Rprofile"))` to edit .Rprofile in HOME.

If multiple rprofiles are present, use `file.edit(".Rprofile")` to edit the project level file.

In the .Rprofile file, a line can be added, `options("turbokit-optionname")`

## boot

Turbokit provides powerful features for starting the package automatically, and reduce redundant pre-analysis tasks such as library calls. However, changing a users' namespace is not always wanted. By default, this options is turned off.

```{r, eval=FALSE}
getOption("turbokit-boot")
```

A message is printed if verbose is set to `TRUE`.

### up

Up, or user preference, is an option that determines the output of several of turbokits' functions. It specifies a maximum of 9 different packages to search when using the construct_complex function. Additionally, it determines the prefix to abbreviations to call a function. To get the current options.

```{r, eval=FALSE}
getOption("turbokit-up")
```

to change its option, there is a function available. See the Custom bindings in the Use turbo section.

```{r, eval=FALSE}
read(package_name)
```

To undo the read functions' doings, another function is shipped with the package, which returns the default option list.

```{r, eval=FALSE}
default()
```

### autoinsert

For Windows users, Turbokit can automatically insert the snippet. By default, this option is set to `TRUE`. If the package `devtools` is not installed, this function will not be available. To check your settings, use

```{r, eval=FALSE}
getOption("turbokit-autoinsert")
```

to disable, use

```{r, eval=FALSE}
options("turbokit-autoinsert" = FALSE)
```

## smartpipe

The smart pipe shortcut chooses between `%>%` and `+` pipe operators based on the context of the text and applies indentation. The option smartpipe changes the behavior of the shortcut where the pipe is inserted. When set to `TRUE`, the pipe is inserted at the end of the line, and the shortcut can thus be called while the cursor is located within brackets of a function. When set to `FALSE`, the operator will be inserted at the cursor location.

The default option is is set to `TRUE`. To check your settings, use

```{r, eval=FALSE}
getOption("turbokit-smartpipe")
```

to disable, use

```{r, eval=FALSE}
options("turbokit-smartpipe" = FALSE)
```

## snippetdir

The directory where snippets are stored is likely to be untouched by many R users. However, when run in a cloud environment or when the directory has been manually changed, users may want to change this option so that turbokit can find the proper location of the snippet file to change.

The default for this is the value returned by an internal `.get_snippets_path()` function. When called from a Windows PC, this will return

    %appdata%/Roaming/RStudio/snippets/r.snippets

For other platform users, it will return

    ~/.config/rstudio/snippets/r.snippets

Users can change this behavior by setting a directory manually in the form of a character vector. If unsure about how to produce this, it is recommended to run

```{r, eval=FALSE}
file.path(file.choose())
```

and navigate to the location of a `r.snippets` file. Other language snippets are not supported by the turbokit package, but do notice - no checks are run on the filepath, make sure you don't accidentally overwrite other language files.

## verbose

Some users may not want to be updated regularly on the changes in behavior of functions. To turn of the chatty behavior of the package,

```{r, eval=FALSE}
getOption("turbokit-verbose")
```

to disable, use

```{r, eval=FALSE}
options("turbokit-verbose" = FALSE)

```

# Turbokit Design Philosophy

How many functions are typed, how many backspaces hit every day, every month, every year, per career, per R user? What is the total redundant hand travel distance?

> Do not Repeat Yourself (Hunt & Davis, 1999)

Once these questions are asked, several problems becomes clear.

1.  Much of a function name is redundant for the developer.

    > It deosn't mttaer in waht oredr the ltteers in a wrod are, hmuans can udnertsand it.

2.  Thinking of code is faster than writing code.

3.  Existing systems are powerful, but unpredictable output or inflexible.

Consider the following piece of legacy code:

```{r BaseR, eval=FALSE}
transform(iris[which(iris$Species == "versicolor"),], new = Sepal.Length*2)
```

In baseR such a call would take roughly 76 keystrokes, given that there are no mistakes. Such code is rarely written these days, as R has seen many years of active development. In recent years, package developers started from scratch, and thought deeply about how code is written, introducing the *tidyverse.*

> The goal of the tidyverse is to discover the minimal set of chunks needed to do data science and have some sense of the priority of the remainder.
>
> -Tidyverse design guide / Tidyverse team

Convergence by the user-base toward acceptance of human-centered, consistent, composable and [inclusive unifying principles](https://design.tidyverse.org/index.html) of the tidyverse and increasingly widespread usage of its [style guide](https://style.tidyverse.org/) is great for reducing redundant loss of resources - making R (and data science!) easier.

The tidyverse packages provide an alternative way to write code in a nice format

```{r, eval=FALSE}
iris %>%
    select(starts_with("Sepal"), Species) %>%
    filter(str_detect(Species, "v")) %>%
    group_by(Species) %>%
    summarise(across(everything(), mean)) %>%
    ggplot(aes(Sepal.Length, Sepal.Width)) +
    geom_density_2d_filled()
```

What the current design philosophy does not do, however, is heavily reduce the keypresses required to arrive at this code. The above example takes roughly 75 keystrokes.

Turbokit is implemented with the following principles in mind.

### 1. Writing code is a bottleneck

Code is written for different stakeholders, including the PC, the developer and the end-user. However, it is the end result that counts, not the methodology to arrive at that result. It does not matter if a user fully types out a function call, how many times a piece of code is rewritten, or if the code itself is the result of a function.

Adults read about 270-300 words per minute, but type only about 40-70 words per minute given *fluency* in a language. Now extend this to non-native speakers and different keyboard layouts or scripts. An example is Chinese, where the *Shift* key is often linked to a change in input styles. Punctuation marks and input often is not identical to what R expects and recognizes, for example the parenthesis ）verses ).

The typing process is inherently prone to errors.

*Example*: `"~s*sw>m*2sd>z*a*e>p>1gv>1syc"`, which is equivalent to

```{r, eval=FALSE}
mydata %>%
  select(starts_with()) %>%
  mutate(new = str_detect()) %>%
  summarise(across(everything(), func)) %>%
  ggplot(aes(param)) +
  geom_violin() +
  scale_y_continuous()
```

### 2. Code input is personal

Every user is different. Typing speed, language fluency, R history, interest et cetera all differ in users. Code input methodology should reflect the different preferences of each user.

> Code input methods should be adaptive to user preferences

### 3. Code input should be predictable

Using style guides and naming conventions are a matter of taste. However, there are similarities in all naming conventions. Often names are chunks by separators such as underscores, capital letter or dots that specify verbs and nouns and some metadata of a function. To memorise which function in which package uses what naming convention is a big hurdle in learning any programming language.

This extends to parameters. Non-standard evaluation (NSE) enables different forms of input. Should a parameter be quoted, unquoted, a list or a vector?

> Code input methods should reduce cognitive load required to use functions by being predictable
